---
title: "Predicting Credit Risk for German Loan Applicants"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
    toc: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.align='center', message = FALSE, warning=FALSE)
```

## Goal

Minimization of risk and maximization of profit on behalf of the bank.

To minimize loss from the bank’s perspective, the bank needs a decision rule regarding who to give approval of the loan and who not to. An applicant’s demographic and socio-economic profiles are considered by loan managers before a decision is taken regarding his/her loan application. A predictive model developed on this data is expected to provide a bank manager guidance for making a decision whether to approve a loan to a prospective applicant based on his/her profiles.

## Load Packages

```{r}
library(data.table)
```


## Load Data

```{r}
credit = fread("german_credit.csv", header="auto")

set.seed(123)
trainIndex <- createDataPartition(iris$Species, p = train_split_prop, 
                                  list = FALSE, 
                                  times = 1)

```

